CF Registry Analysis - ROC

### 2024-08-07 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
ridiculously inefficient ROC code but it's working

---
title: "CF Registry Analysis - ROC"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    fig-width: 8
    fig-height: 8
    page-layout: full
bibliography: /Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Miscellaneous/zotero.bib
csl: /Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Miscellaneous/american-medical-association.csl
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(gtsummary)
library(gt)
library(pROC)
switch(Sys.info()[["sysname"]],
  Windows = {
    home_dir <- "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry"
  },
  Darwin = {
    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry"
  }
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

## Three groups, two cutoffs

@@

### Peds and adults

@@

```{r}
#| message: false
r <- roc(ogtt ~ hgba1c, data = ngt_agt)
plot(r, print.auc = T, print.thres = T)
gt(coords(r, ret = metrics)) %>% fmt_number(n_sigfig = 3)
```

@@

```{r}
#| message: false
r <- roc(ogtt ~ hgba1c, data = agt_cfrd)
plot(r, print.auc = T, print.thres = T)
gt(coords(r, ret = metrics)) %>% fmt_number(n_sigfig = 3)
```

### Peds

@@

```{r}
#| message: false
r <- roc(ogtt ~ hgba1c, data = peds_ngt_agt)
plot(r, print.auc = T, print.thres = T)
gt(coords(r, ret = metrics)) %>% fmt_number(n_sigfig = 3)
```

```{r}
#| message: false
r <- roc(ogtt ~ hgba1c, data = peds_agt_cfrd)
plot(r, print.auc = T, print.thres = T)
gt(coords(r, ret = metrics)) %>% fmt_number(n_sigfig = 3)
```

### Adults

@@

```{r}
#| message: false
r <- roc(ogtt ~ hgba1c, data = adult_ngt_agt)
plot(r, print.auc = T, print.thres = T)
gt(coords(r, ret = metrics)) %>% fmt_number(n_sigfig = 3)
```

```{r}
#| message: false
r <- roc(ogtt ~ hgba1c, data = adult_agt_cfrd)
plot(r, print.auc = T, print.thres = T)
gt(coords(r, ret = metrics)) %>% fmt_number(n_sigfig = 3)
```

## Two groups, one cutoffs

@@

```{r}
#| message: false
r <- roc(agt ~ hgba1c, data = roc_df)
plot(r, print.auc = T, print.thres = T)
gt(coords(r, ret = metrics)) %>% fmt_number(n_sigfig = 3)
```

@@

```{r}
#| message: false
r <- roc(cfrd ~ hgba1c, data = roc_df)
plot(r, print.auc = T, print.thres = T)
gt(coords(r, ret = metrics)) %>% fmt_number(n_sigfig = 3)
```

### Peds

@@

```{r}
#| message: false
r <- roc(agt ~ hgba1c, data = peds_roc)
plot(r, print.auc = T, print.thres = T)
gt(coords(r, ret = metrics)) %>% fmt_number(n_sigfig = 3)
```

```{r}
#| message: false
r <- roc(cfrd ~ hgba1c, data = peds_roc)
plot(r, print.auc = T, print.thres = T)
gt(coords(r, ret = metrics)) %>% fmt_number(n_sigfig = 3)
```

### Adults

@@

```{r}
#| message: false
r <- roc(agt ~ hgba1c, data = adult_roc)
plot(r, print.auc = T, print.thres = T)
gt(coords(r, ret = metrics)) %>% fmt_number(n_sigfig = 3)
```

```{r}
#| message: false
r <- roc(cfrd ~ hgba1c, data = adult_roc)
plot(r, print.auc = T, print.thres = T)
gt(coords(r, ret = metrics)) %>% fmt_number(n_sigfig = 3)
```

>>>>

---
title: "CF Registry Analysis - ROC"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    fig-width: 8
    fig-height: 8
    page-layout: full
bibliography: /home/tim/Documents/Miscellaneous/zotero.bib
csl: /home/tim/GitHub/styles/american-medical-association.csl
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(gtsummary)
library(gt)
library(pROC)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry",
  "Linux" = "/home/tim/OneDrive/Vigers/CF/Christine Chan/CF Registry"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/GitHub",
  "Windows" = "C:/Users/Tim/GitHub",
  "Linux" = "/home/tim/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

For each ROC curve, we performed 2000 bootstrap replicates. Here the median best threshold (based on Youden index), sensitivity, and specificity are presented along with the bootstrapped 95% confidence intervals. Column names are printed in terms of percentiles, so `threshold.2.5.` indicates the lower limit of the confidence interval, `threshold.50.` indicates the median best threshold, and `threshold.97.5.` indicates the upper limit of the confidence interval.

## NGT vs. AGT and AGT vs. CFRD

@@

### Full cohort

@@

```{r}
#| message: false
r <- roc(ogtt ~ hgba1c, data = ngt_agt)
plot(r, print.auc = T, print.thres = T)
gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
```

@@

```{r}
#| message: false
r <- roc(ogtt ~ hgba1c, data = agt_cfrd)
plot(r, print.auc = T, print.thres = T)
gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
```

### Peds vs. Adults

#### Peds

@@

##### NGT vs. AGT

```{r}
#| message: false
r <- roc(ogtt ~ hgba1c, data = peds_ngt_agt)
plot(r, print.auc = T, print.thres = T)
gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
```

##### AGT vs. CFRD

```{r}
#| message: false
r <- roc(ogtt ~ hgba1c, data = peds_agt_cfrd)
plot(r, print.auc = T, print.thres = T)
gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
```

#### Adults

@@

##### NGT vs. AGT

```{r}
#| message: false
r <- roc(ogtt ~ hgba1c, data = adult_ngt_agt)
plot(r, print.auc = T, print.thres = T)
gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
```

##### AGT vs. CFRD

```{r}
#| message: false
r <- roc(ogtt ~ hgba1c, data = adult_agt_cfrd)
plot(r, print.auc = T, print.thres = T)
gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
```

### By race

#### White

```{r}
white_ngt_agt <- ngt_agt %>% filter(Race == "White")
white_agt_cfrd <- agt_cfrd %>% filter(Race == "White")
```

##### NGT vs. AGT

```{r}
#| message: false
r <- roc(ogtt ~ hgba1c, data = white_ngt_agt)
plot(r, print.auc = T, print.thres = T)
gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
```

##### AGT vs. CFRD

```{r}
#| message: false
r <- roc(ogtt ~ hgba1c, data = white_agt_cfrd)
plot(r, print.auc = T, print.thres = T)
gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
```

#### Black or African American

```{r}
black_ngt_agt <- ngt_agt %>% filter(Race == "Black or African American")
black_agt_cfrd <- agt_cfrd %>% filter(Race == "Black or African American")
```

##### NGT vs. AGT

```{r}
#| message: false
r <- roc(ogtt ~ hgba1c, data = black_ngt_agt)
plot(r, print.auc = T, print.thres = T)
gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
```

##### AGT vs. CFRD

Bootstrap replicates found more than one "best" threshold, so the non-bootstrapped threshold is reported first, and a random bootstrapped "best" interval second.

```{r}
#| message: false
r <- roc(ogtt ~ hgba1c, data = black_agt_cfrd)
plot(r, print.auc = T, print.thres = T)
gt(data.frame(coords(r, "best"))) %>%
  fmt_number(n_sigfig = 3)
gt(data.frame(ci.coords(r, "best", best.policy = "random"))) %>%
  fmt_number(n_sigfig = 3)
```

#### More than one race

```{r}
more_ngt_agt <- ngt_agt %>% filter(Race == "More than one race")
more_agt_cfrd <- agt_cfrd %>% filter(Race == "More than one race")
```

##### NGT vs. AGT

```{r}
#| message: false
r <- roc(ogtt ~ hgba1c, data = more_ngt_agt)
plot(r, print.auc = T, print.thres = T)
gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
```

##### AGT vs. CFRD

```{r}
#| message: false
r <- roc(ogtt ~ hgba1c, data = more_agt_cfrd)
plot(r, print.auc = T, print.thres = T)
gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
```

#### Other race

```{r}
other_ngt_agt <- ngt_agt %>% filter(Race == "Other race")
other_agt_cfrd <- agt_cfrd %>% filter(Race == "Other race")
```

##### NGT vs. AGT

Bootstrap replicates found more than one "best" threshold, so the non-bootstrapped threshold is reported first, and a random bootstrapped "best" interval second.

```{r}
#| message: false
r <- roc(ogtt ~ hgba1c, data = other_ngt_agt)
plot(r, print.auc = T, print.thres = T)
gt(data.frame(coords(r, "best"))) %>%
  fmt_number(n_sigfig = 3)
gt(data.frame(ci.coords(r, "best", best.policy = "random"))) %>%
  fmt_number(n_sigfig = 3)
```

##### AGT vs. CFRD

Bootstrap replicates found more than one "best" threshold, so the non-bootstrapped threshold is reported first, and a random bootstrapped "best" interval second.

```{r}
#| message: false
r <- roc(ogtt ~ hgba1c, data = other_agt_cfrd)
plot(r, print.auc = T, print.thres = T)
gt(data.frame(coords(r, "best"))) %>%
  fmt_number(n_sigfig = 3)
gt(data.frame(ci.coords(r, "best", best.policy = "random"))) %>%
  fmt_number(n_sigfig = 3)
```

## Two hour OGTT >= 140 or >= 200

@@

```{r}
#| message: false
r <- roc(agt ~ hgba1c, data = roc_df)
plot(r, print.auc = T, print.thres = T)
gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
```

@@

```{r}
#| message: false
r <- roc(cfrd ~ hgba1c, data = roc_df)
plot(r, print.auc = T, print.thres = T)
gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
```

### Peds vs. Adults

#### Peds

@@

##### $\geq$ 140

```{r}
#| message: false
r <- roc(agt ~ hgba1c, data = peds_roc)
plot(r, print.auc = T, print.thres = T)
gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
```

##### $\geq$ 200

```{r}
#| message: false
r <- roc(cfrd ~ hgba1c, data = peds_roc)
plot(r, print.auc = T, print.thres = T)
gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
```

#### Adults

@@

##### $\geq$ 140

```{r}
#| message: false
r <- roc(agt ~ hgba1c, data = adult_roc)
plot(r, print.auc = T, print.thres = T)
gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
```

##### $\geq$ 200

```{r}
#| message: false
r <- roc(cfrd ~ hgba1c, data = adult_roc)
plot(r, print.auc = T, print.thres = T)
gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
```

### By race

#### White

```{r}
white_roc <- roc_df %>% filter(Race == "White")
```

##### $\geq$ 140

```{r}
#| message: false
r <- roc(agt ~ hgba1c, data = white_roc)
plot(r, print.auc = T, print.thres = T)
gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
```

##### $\geq$ 200

```{r}
#| message: false
r <- roc(cfrd ~ hgba1c, data = white_roc)
plot(r, print.auc = T, print.thres = T)
gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
```

#### Black or African American

```{r}
black_roc <- roc_df %>% filter(Race == "Black or African American")
```

##### $\geq$ 140

```{r}
#| message: false
r <- roc(agt ~ hgba1c, data = black_roc)
plot(r, print.auc = T, print.thres = T)
gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
```

##### $\geq$ 200

```{r}
#| message: false
r <- roc(cfrd ~ hgba1c, data = black_roc)
plot(r, print.auc = T, print.thres = T)
gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
```

#### More than one race

```{r}
more_roc <- roc_df %>% filter(Race == "More than one race")
```

##### $\geq$ 140

```{r}
#| message: false
r <- roc(agt ~ hgba1c, data = more_roc)
plot(r, print.auc = T, print.thres = T)
gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
```

##### $\geq$ 200

```{r}
#| message: false
r <- roc(cfrd ~ hgba1c, data = more_roc)
plot(r, print.auc = T, print.thres = T)
gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
```

#### Other race

```{r}
other_roc <- roc_df %>% filter(Race == "Other race")
```

##### $\geq$ 140

Bootstrap replicates found more than one "best" threshold, so the non-bootstrapped threshold is reported first, and a random bootstrapped "best" interval second.

```{r}
#| message: false
r <- roc(agt ~ hgba1c, data = other_roc)
plot(r, print.auc = T, print.thres = T)
gt(data.frame(coords(r, "best"))) %>%
  fmt_number(n_sigfig = 3)
gt(data.frame(ci.coords(r, "best", best.policy = "random"))) %>%
  fmt_number(n_sigfig = 3)
```

##### $\geq$ 200

Bootstrap replicates found more than one "best" threshold, so the non-bootstrapped threshold is reported first, and a random bootstrapped "best" interval second.

```{r}
#| message: false
r <- roc(cfrd ~ hgba1c, data = other_roc)
plot(r, print.auc = T, print.thres = T)
gt(data.frame(coords(r, "best"))) %>%
  fmt_number(n_sigfig = 3)
gt(data.frame(ci.coords(r, "best", best.policy = "random"))) %>%
  fmt_number(n_sigfig = 3)
```
